Davos: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-09-23 paper/main.tex
edits through end of section 2

%- .configure()
\item \texttt{configure(**kwargs)}: This function provides an alternate way of assigning values to the writeable attributes listed in Section \ref{subsec:config} and can be used to configure multiple options at once (see Sec.~\ref{sec:illustrative-example} for example usage). The function accepts attribute names as keyword-only arguments, to which their desired values are passed. If any of the options passed are incompatible (e.g., both \texttt{confirm\_install=True} and \texttt{noninteractive=True} are passed) or assignment to any of the specified attributes fails for any reason, none of the given options will be modified.

%- .get_project()
\item \texttt{get\_project(project\_name, create=False)}: This function can be passed the name of a Davos project (\texttt{project\_name}) to get the \texttt{Project} or \texttt{AbstractProject} instance representing it. If no project with the given name exists and \texttt{create=False} (the default), the function will return \texttt{None}; if \texttt{create=True}, a project with the given name will instead be created and returned.

%- .prune_projects()
\item \texttt{prune\_projects(yes=False)}: This function allows users to quickly ``clean up'' their local Davos projects by deleting notebook-specific projects whose corresponding notebooks no longer exist (i.e., \texttt{AbstractProject}s). As with standard virtual environments, periodically removing large package directories can be useful for reclaiming disk space once they are no longer needed. By default, this function will interactively display a list of all unused projects and allow the user to choose whether or not to delete each one. Alternatively, passing \texttt{yes=True} will immediately remove all unused projects without requiring confirmation. Note that if Davos's non-interactive mode is enabled (i.e., \texttt{davos.noninteractive} has been set to \texttt{True}), \texttt{yes=True} \textit{must} be explicitly passed in order to delete projects. This serves as a safeguard against accidentally deleting projects, since non-interactive mode disables all user input and confirmation.  Also note that this function will not delete notebook-agnostic projects (i.e., manually created projects whose names are not notebook filepaths), as they are not linked to a specific notebook whose existence can be checked automatically. However, any project can be deleted individually via its \texttt{Project} object's \texttt{.remove()} method.

%- .require_pip()
%\item \texttt{require\_pip(version\_spec, warn=False, extra\_msg=None, prereleases=None)}:
%- .require_python()
\item \texttt{require\_python(version\_spec, warn=False, extra\_msg=None, prereleases=None)}:
%- .use_default_project()
\item \texttt{use\_default\_project()}:
\end{itemize}

@@

>>>>

\item \texttt{configure(**kwargs)}: This function provides an alternate way of assigning values to the writeable attributes listed in Section \ref{subsec:config} and can be used to configure multiple options at once (see Sec.~\ref{sec:illustrative-example} for example usage). The function accepts attribute names as keyword-only arguments to which their desired values are passed. If any of the options passed are incompatible (e.g., both \texttt{confirm\_install=True} and \texttt{noninteractive=True} are passed) or assignment to any of the specified attributes fails for any reason, none of the given options will be modified.

\item \texttt{get\_project(project\_name, create=False)}: This function can be passed the name of a Davos project (\texttt{project\_name}) to get the \texttt{Project} or \texttt{AbstractProject} instance representing it. The optional \texttt{create} argument determines the function's behavior when no project with the given name exists: if \texttt{create=False} (the default), the function will return \texttt{None}; if \texttt{create=True}, a project with the given name will be created and returned.

%If no project with the given name exists and \texttt{create=False} (the default), the function will return \texttt{None}; if \texttt{create=True}, a project with the given name will instead be created and returned.

\item \texttt{prune\_projects(yes=False)}: This function allows users to quickly ``clean up'' their local Davos projects by deleting notebook-specific projects whose corresponding notebooks no longer exist (i.e., \texttt{AbstractProject}s). As with standard virtual environments, periodically removing unused project directories can be useful for reclaiming disk space from dependencies of code that is no longer in use. By default, this function will interactively display a list of all unused projects and allow the user to choose whether or not to delete each one. Alternatively, passing \texttt{yes=True} will immediately remove all unused projects without prompting for confirmation. Note that if Davos's non-interactive mode is enabled (see Sec.~\ref{subsec:config}), \texttt{yes=True} must be explicitly passed, otherwise the function will raise an exception. This serves as a safeguard against accidentally deleting projects since non-interactive mode disables all user input and confirmation.  Also note that this function will not delete notebook-agnostic projects (i.e., manually created projects whose names are not notebook filepaths), as they are not linked to specific notebooks whose existence determines whether or not they are still needed. These (and any) projects may be deleted individually by calling their \texttt{Project} objects' \texttt{.remove()} method.

\item \texttt{require\_python(version\_spec, warn=False, extra\_msg=None, prereleases=None)}: Through \texttt{smuggle} statements and onion comments, Davos can automatically ensure that all Python packages needed to run a notebook are installed, and that the same versions of those packages are used no matter when or by whom the notebook is run. However, because Davos operates at runtime, one thing it cannot do automatically is install and switch to a specific version of Python itself. 
Distributing shared code along with a particular Python version for running it 
%Ensuring that a particular Python version is used to run shared code typically
%Ensuring that shared code is run with a particular Python version typically 
requires a heavier-weight solution,
%than Davos (or a virtual environment), 
such as a Conda environment or Docker container (see Fig.~\ref{fig:code-sharing}). 
Yet a Davos-enhanced notebook may still \texttt{smuggle} certain packages that depend on users having a specific Python version or range of versions (e.g., even just within the standard library, the \texttt{dataclass} module was first added in Python 3.7 \cite{Smit17} and at least 19 modules are slated for removal in Python 3.13 \cite{HeimCann19}). The \texttt{davos.require\_python()} function can be added to the top of a Davos-enhanced notebook to communicate to users that the notebook's code should be run with a specific or constrained Python version (see Sec.~\ref{sec:illustrative-example} for example usage). The function may be passed a version string (e.g., \texttt{"3.10.5"}) or any valid version specifier \cite{CoghStuf13} (e.g., \texttt{"\raisebox{0.5ex}{\texttildelow}=3.11"}, \texttt{">=3.9;<3.12"}, etc.) and will raise an exception if the user's Python version is incompatible. Alternatively, a ``soft'' or suggested constraint can be specified by passing \texttt{warn=True} to issue a warning rather than raise an error. Additional information can be added to the default error/warning message (e.g., the specific reason for this requirement) via the \texttt{extra\_msg} argument, and the optional \texttt{prereleases} argument can be used to explicitly allow (\texttt{True}) or disallow (\texttt{False}) pre-release versions (by default, the policy is determined by the value of \texttt{version\_spec}).

\item \texttt{use\_default\_project()}: By default, each Davos-enhanced notebook will create and use a notebook-specific project named for the absolute path to the notebook file. If a user manually changes the project used by the current notebook (i.e., by setting the value of the \texttt{davos.project} attribute; see Sec.~\ref{subsec:config}), this function can be called to reset \texttt{davos.project} to its default value and switch back to the original default project. See Section \ref{subsec:projects} for more information about Davos projects and Section \ref{sec:illustrative-example} for an illustrative example.
%If the user manually configures a notebook to use a different project (i.e., by setting the value of the \texttt{davos.project} attribute; see Sec.~\ref{subsec:config}), this function can be called to 
%This function can be called to reset the \texttt{davos.project} attribute (i.e., the project used by the current notebook; see Sec.~\ref{subsec:config}) to its default value if it has been manually set to a different value. The default project for a Davos-enhanced notebook will be a notebook-specific project named for the notebook's absolute path (see Sec.~\ref{subsec:projects}), with two (uncommon) exceptions. First, First, if Davos cannot determine \_\_\_. Second, Davos offers experimental support for use in IPython terminal shells. All terminal shell instances share a common project named \texttt{"ipython-shell"}.

@@

\bigskip\bigskip\textcolor{red}{\textbf{========== TODO: finish editing from here to end ==========}}\bigskip\bigskip
